Manifold fitting by ridge estimation: a subspace-constrained approach

### 2021-02-25 Ridge_Estimation_main/Ridge_LKDE_v2-edited.tex
zzh

Today, the process of dealing with high-dimensional data requires the exponential consumption of computational resources, and is therefore costly but effective. Yet, even so, challenges remain, as the high-dimensional data tend to lie near a low-dimensional sub-manifold of the ambient space, which is a phenomenon termed ``the curse of dimensionality'' by \cite{fefferman2018fitting}.  Although, the geometric foundation of statistical inference in nonlinear regression was brought to attention of the statistical community as early as the 1980s by \cite{Kass1989}, the role of geometry was not primarily explored back then. At present, there are numerous works concerning manifold learning, with different areas of focus, such as dimension reduction focusing on approximating the embedding \citep{roweis2000nonlinear,zhang2004principal,donoho2003hessian,zha2007} and manifold approximation via fitting the unknown manifold \citep{genovese2012manifold,chen2015asymptotic,genovese2014nonparametric}. In another seemingly-related direction, there have been works on statistical methods on manifolds over the past decades, centering on finding the mode of distribution \citep{huckemann2010intrinsic,chen2016nonparametric,chen2016comprehensive} and searching for the principal components of manifold \citep{hauberg2015principal,huckemann2006principal,hastie1989principal}.

There has also been some theoretical research into manifold learning using hypothesis testing. For instance, \cite{fefferman2018fitting} provided a condition and proved that the produced putative manifold ${\mathcal M}_o$'s Hausdorff distance to $\mathcal M$ is small, and its reach is not much smaller than that of $\mathcal M$. Mohammed and Narayanan proposed that the output manifold $\mathcal M$ can be defined by the set of points where the gradient of the approximate squared distance function (asdf) is orthogonal to the subspace spanned by the largest $D-d$ eigenvectors of the Hessian of the asdf \citep{mohammed2017manifold}.

@@

Manifold-fitting algorithms intend to recover the low-dimensional manifold in the ambient space by supposing that the observation data is generated from the manifold with some noise added. There are many classical algorithms to find the lower-dimensional manifold, such as the principal flow \citep{panaretos2014principal} and the local principal curves and surfaces based on kernel-density estimation \citep{ozertem2011locally}. The principal flow attempts to fit a curve moving along the maximal variation of the data, subject to a smoothness constraint. The solution to the principal flow is reduced to an ODE problem via the Euler-Lagrange method.  \cite{ozertem2011locally} redefined the principal curves and surfaces in terms of the gradient and Hessian of the probability density estimate. They also provided an algorithm based on the classical KDE. \cite{jung2012analysis} propose a general framework to fit the high-dimensional data with principal nested spheres. In addition, there is a large amount of literature about searching for hidden structure with an application in the point clouds, nebular hypothesis, earthquake locations, and object-oriented objects \citep{adams2011morse,davenport2010joint,klemela2009smoothing,book}.

@@

\begin{figure}[h] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=\linewidth]{field_demo.eps} 
   \caption{Illustration of the vector field. Left: using partial nearest samples Right: using all samples}
   \label{fig:example:1}
\end{figure}

The effect of improving the quality of $C(x)$ by neglecting samples with a large scale can be seen in the vector field in Figure \ref{fig:example:1}. The vector field is constructed from the dominant eigenvector of the weighted covariance matrix $C(x)$. In the left diagram, the set $S$ is the $k$-nearest neighborhood of $x$, and in the right diagram the set $S$ represents all the samples. From the figure, we know that by selecting the nearest samples we can get a weighted covariance matrix to approximate the vector field better.

Based on the above intuition, we show that constructing the {\it l}-SCRE by neglecting the far-away samples will help to reduce the point-wise bias of the derivatives. The reduced-bias phenomenon will further lead us to a ridge that is `closer' than that obtained by using all the samples.
 \subsection{Main Contributions}
Our contributions consist mainly of the following:
\begin{itemize}
\item[1.] We propose to use the nearest samples to construct a local version of the SCRE method, and in turn use that to construct the ridge-estimator function. The proposed {\it l}-SCRE  can obtain a ridge that is closer to the original manifold.
\item[2.] We provide the theoretical result of the component-wise bias and variance corresponding to the {\it l}-SCRE and prove that, under mild conditions, the local kernel function can indeed reduce the bias of derivatives more effectively than the KDE-based approaches.
\item[3.] By building the connection Hessian of the estimated density function with the empirical covariance matrix, we show the necessity of rank-one modification to obtain a stable tangent space estimator, which further improves the performance of the {\it l}-SCRE.
%\item[1.] We develop the connection between the two common methods (kernel-density estimation and tangent-space estimation) to define the fitted manifold.
%\item[2.] By analyzing the eigenspace of the Hessian matrix $H(x)$ of the kernel-density estimation function, we show that the normal space is not always generated by the principal eigenvectors of $H(x)$. To solve this problem, we split the $H(x)$ into two parts, $A(x)$ and $B(x)$. By estimating the normal space from $B(x)$, we can get a stable estimate of the normal space. The numerical experiment also demonstrates the effectiveness of our method.
%\item[3.] Our method can not only move an outlier $x_i$ onto the manifold, but also restrict the path of the convergence to be in the normal space of the intended targeting point $c(x)$.
%\item[4.] We also provide a faster algorithm to cope with the vanishing-gradient phenomenon for the kernel-density function.
\end{itemize}
Finally, we provide numerical examples to show the effectiveness of our strategy, and also show that the vector field corresponding to our method is smoother and more appropriate for the manifold-estimation problem.
\subsection{Notations}
The most frequently used symbols and notations are listed in Table 1.%\ref{symbols}.

\begin{table}[h]\label{symbols}
\caption{Symbols and Notations}
\resizebox{14cm}{!}{
\begin{tabular}{c|c}\hline
Symbols & Meaning and Explanation\\\hline
$d$,$D$ & The dimension of the manifold and the dimension of the ambient space\\
${\rm E},{\rm Var}$ & The expectation and the variance operator\\
$\mathcal M,\hat{\mathcal G}$ & The true unknown manifold and the ridge estimator (discrete set)\\ 
$\mathcal M_{\hat{G}}$ & The projection of the ridge estimator $\hat{\mathcal G}$ onto $\mathcal M$\\
$p_n(x),g_n(x),H(x)$& The smooth density function, its gradient, and the Hessian Matrix\\
$\hat{p}_{r,h}(x), g_{\hat{p}_{r,h}}(x),H_{\hat{p}_{r,h}}(x)$& The local KDE estimator, its gradient, and the Hessian Matrix\\
 $\Pi,\Pi^c$ & The projection onto the top $d$ eigenvectors and its orthogonal complement\\
 $f$ & The nonlinear increasing, nonnegative, and concave function\\ 
 $J(x)$ & The semi-positive definite matrix, which is a key component of the Hessian\\  
 {$C(x)$, $H(x)$} &  {The weighted sample covariance matrix and the Hessian matrix at $x$}\\
 $\Pi_H$ & The projection matrix constructed by the top $d$ eigenvectors of $H$ \\ 
 ${\mathcal D}_r(x_0)$ & The $D$-dimensional ball centered at $x_0$ with a radius of $r$\\
 ${\cal N}(x, r)$ & The samples in the neighborhood of $x$ with a radius of $r$ \\
 \hline 
\end{tabular}}
\end{table}

@@

We adopt the classical kernel-density estimation approach to handle our manifold-fitting problem, as the majority of previous works do. From the kernel-density perspective, the manifold can be viewed as points that satisfy some specific ridge conditions of the estimated density function.
\subsection{Ridge Definition}
There are two main definitions of a ridge, both of which involve describing the eigenspace of the Hessian and the gradient.
\begin{definition}\label{RR}
A point $x$ is on the $d$-dimensional ridge $R$ of its probability density function when the gradient $g(x)$ is orthogonal to at least $D-d$ eigenvectors of $H(x)$ and the corresponding $D-d$ eigenvalues are negative. See \cite{ozertem2011locally}
\end{definition}
This definition gives us only a condition to test whether a point $x$ lies on the $d$-dimensional ridge. Given any $x$, it does not necessarily tell us how to pull it onto the $d$-dimensional ridge. There are $D-d$ eigenvectors of $H(x)$ that are orthogonal to the gradient $g(x)$, but it does not tell us the eigenvalues of $H(x)$ to which the eigenvectors correspond.

@@

\subsection{Ridge Estimation and Error Measurement}
In practice, we use the ridge $R$ as an approximation of the unknown manifold $\mathcal M$ numerically. However, the definition of a ridge in \eqref{ridgeDF} is arguably too descriptive and unrealistic. We can only judge whether a point resides on the ridge, but the definition provides an explicit continuous function to describe the ridge. In reality, starting from a random sample given in advance, we will need an algorithm to push the data points onto the ridge so that the resultant points satisfy the ridge definition. To be specific, what we obtain is a set of modified discrete samples, $\hat{\mathcal G}$, satisfying the ridge definition. To this end, we may further specify the ridge estimator $\hat{\cal G}$ in \eqref{Estimator} by
\[
\hat{\mathcal G} = {\rm Algorithm}(\{x_i, i=1:n\},  \hat{R})
\]
where the estimated $\hat{R}$ can be $\hat{R}_h$ or $\hat{R}_{r,h}$. Note that the resultant $\hat{\mathcal G}$ relies on the initial data points $\{x_i\}$, the estimated ridge $\hat{R}$, and the actual implementation (algorithm) involved in determining how to process the data $\{x_i\}$ by moving them onto the defined ridge. Obviously, we know $\hat{\mathcal G} \subset \hat{R}$.
\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=3.5in]{introduction_demo.eps} 
      \caption{Illustration of the $\hat{\mathcal G},\mathcal M, \mathcal M_{\hat{\mathcal G}}$ and the Hausdorff and average margin using the example of a toy circle. The blue diamond markers represent the observation points. The blue dots represent the computed points that satisfy the ridge condition, i.e. $\hat{\mathcal G}$. The red dots represent the projection of $\hat{\mathcal G}$ onto $\mathcal M$, denoted by $\cal M_{\hat{G}}$. }
   \label{fig:example:2}
\end{figure}

@@

To diminish the random factor in the distance measurement between $\hat{\mathcal G}$ and $\mathcal M$, we extract a subset from $\mathcal M$, ${\mathcal M}_{\hat{\mathcal G}}=\pi_{\mathcal M}(\hat{\mathcal G})$, and consider the Hausdorff distance between ${\hat{\mathcal G}}$ and ${\mathcal M}_{\hat{\mathcal G}}$ (see Figure \ref{fig:example:2}). Since the projection $\pi_{\mathcal M}(\hat{\mathcal G})$ from $\hat{\mathcal G}$ to $\mathcal M_{\hat{G}}$ is onto, the Hausdorff distance equals the quasi-Hausdorff distance

@@

Lemma \ref{margin} builds the bridge between the geometric quantities and the analytic differential quantities. Through this lemma, we can easily convert the problem of searching for a `closer' ridge to the problem of searching for a `better' approximation of the Hessian and gradient. As illustrated in Figure \ref{fig:example:3}, this conversion arguably give our work a stronger foundation and greater versatility.

@@

   \label{fig:example:3}

@@

\begin{example}
When we approximate $\hat{p}(x), g_{\hat{p}}(x), H_{\hat{p}}(x)$ using the kernel-density function as $\hat{p}_h(x), g_{\hat{p}_h}(x), H_{\hat{p}_h}(x)$, %where the subscript $n$ means the functions are related with the samples and the sample size $n$.
the pairwise error for the derivatives yields the following result:
\[
\begin{aligned}
& ( |g_{\hat{p}_h}(x)-g_p(x)|)_i = O(h^2) + O_p(\sqrt{\frac{1}{{{n h^{D+2}}}}}),\\
& ( |H_{\hat{p}_h}(x)-H_p(x)| )_{ij} = O(h^2) + O_p(\sqrt{\frac{1}{{{n h^{D+4}}}}}).
\end{aligned}
\]
The empirical-process theory gives the uniform error as
\[
\begin{aligned}
& \sup_x( |g_{\hat{p}_h}(x)-g_p(x)|)_i = O(h^2) + O_p(\sqrt{\frac{\log n}{{{n h^{D+2}}}}}),\\
& \sup_x( |H_{\hat{p}_h}(x)-H_p(x)| )_{ij} = O(h^2) + O_p(\sqrt{\frac{\log n}{{{n h^{D+4}}}}}).
\end{aligned}
\]
The error rates for the pairwise and uniform differ only in the stochastic term, with a scale of $\sqrt{\log n}$. See \cite{chen2017tutorial,JMLR:v17:ariascastro16a, genovese2014nonparametric}.
\end{example}
%From \eqref{Hausmargin}, we can conclude that the Hausdorff distance of ${\mathcal G}$ and ${\mathcal M}_{\mathcal G}$ is controlled by the uniform approximation error of the Hessian and gradient, as $\sup_x( \|H_n(x)-H(x)\|_2+\|g_n(x)-g(x)\|_2$. The uniform derivatives rate is proved to be $L_{\infty}$, and the norm as , where the difference is the result of introducing $\log n$. This result is shown in \cite{JMLR:v17:ariascastro16a, genovese2014nonparametric}.
%Clearly, the Hausdorff distance between is the largest gap. As shown in \cite{genovese2014nonparametric}, the Hausdorff distance between the manifold and the ridge yields a uniform bound 
%${\rm Haus}({\mathcal G}, {\mathcal M}_{\mathcal G}) = O(h^2) + O_p((\frac{\log n}{{{n h^{d+4}}}})^{1/2})$. When $h={(\frac{\log n}{n}})^{1/(d+8)}$, the optimal minimum achieved is $O_p((\frac{\log n}{n})^{2/(d+8)})$. This optimum is the estimated convergent speed for the Hausdorff distance with the process of increasing the sample size $n$ and simultaneously decreasing $h$. However, in this paper, we mainly focus on how to get a closer ridge under the current observations, i.e. the number of samples is fixed.

@@

From Lemma \ref{improve_lemma}, we know that by reducing the bias with a scalar of $\ell$ we will have the optimal minimum solution reduced with a scalar of $\ell^{\frac{D+m}{D+m+4}}$. It should be noted that, with the increase in the dimension $D$, we will obtain a smaller $\ell^{\frac{D+m}{D+m+4}}$. Lemma \ref{improve_lemma} proves that we can improve the distance between the ridges by reducing the derivatives' bias corresponding to the density function. As before, the details can be found in the supplementary material.%appendix \ref{minimum_proof}.

@@

\section{Numerical Experiments}\label{experiments}
In this section, we will show the effectiveness of our method compared with the classical subspace-constrained methods for manifold fitting. The comparison is based on a uniform frame for all methods.
Overall, the manifold algorithm can be split into two parts. For the first part, for any point ${x}$ lying outside the manifold, construct an attraction force starting from ${x}$ and ending with some point (often represented by the observations) on the manifold. For the second part, estimate the normal space and use the projection to modify the attraction force such that $x$'s moving trajectory resembles the projected trajectory as much as possible. We briefly summarize the main methods considered for comparison below and we refer the readers to the Appendix \ref{Uniform_Framework}  for the implementation details (B1-B2) and the Ridge-searching Algorithm for SCRE (B3).

{\bf Implementation:} {\rm SCRE}, {\it l}-SCRE, MFIT-i \citep{fefferman2018fitting}, and MFIT-ii \citep{yao2019manifold}. The MATLAB codes together with all numerical examples used in this paper are available on \url{https://zhigang-yao.github.io/research.html} (for {\rm SCRE}, {\it l}-SCRE and MFIT-ii). We have implemented MFIT-i,  since the authors of \cite{fefferman2018fitting} have not provided implementation due to the nature of their work has been purely abstract.

\subsection{Ridge Criteria}
The criterion for a point on a ridge is that it must satisfy the ridge condition. The ridge condition indicates that, when $x$ is on the ridge, the attraction vector $F(x)$ resides in the tangent space. The tangent space is usually obtained from the top d eigenvectors of the covariance matrix. Conversely, if the ridge condition is not met at $x$, we can also expect the angle between the attraction vector $F(x)$ and the estimated eigenspace to be large, i.e. a small cos value.

@@

   \begin{figure}[h] %  figure placement: here, top, bottom, or page
    \includegraphics[width=\linewidth]{vectorfield2.eps} 
    % \setcaptionwidth{6in}
    \caption{Illustration of the vector field corresponding to the eigenvector of the second eigenvalue with respect to three difference covariance matrices }
    \label{fig:vectorfield}
 \end{figure}
In Figure \ref{fig:vectorfield}, we plot the vector field generated by the eigenvector corresponding to the smallest eigenvalue. This vector field is the geometric interpretation of the projection $\Pi^{\perp}$ in the two-dimensional space. Because the constrained subspace in the two-dimensional space is one-dimensional, the vector field coincides with the trajectory of the outliers moving onto the ridge. We can conclude from the above figures that the vector field corresponding to {\it l}-SCRE is more smooth and points approximately to the projection onto the manifold, while the vector field of SCRE is not well behaved for samples not close to the manifold.
\subsection{Synthetic Data Set}
First, we provide a one-dimensional ring example to show the effectiveness of our method compared with SCRE. Because of the simplicity of the circle structure, it is easy to compute the real projection onto it. We assume that our hidden manifold ${\mathcal M}$ is the one-dimensional circle embedded in a two-dimensional space with a radius of $1$ and center of $(0,0)$.

@@

For each point $\hat{x}$ belonging to ${\cal \hat{G}}$ or ${\cal \hat{G}}_\ell$, we can define the projection of $\hat{x}$ onto the real manifold $\mathcal M$ as $\pi_{\mathcal M}(\hat{x}) = \arg\min_{y\in \mathcal M} \|y-\hat{x}\|_2$. In the case of the special two-dimensional circle, the projection has an explicit form, $\pi_{\mathcal M}(\hat{x}) = \hat{x}/\|\hat{x}\|$. The average margin from $\cal \hat{G}$ or ${ \cal \hat{G}}_\ell$ to $\mathcal M$ is defined as
\[
{\rm Margin}({ \cal \hat{G}, \mathcal M}) =\frac{1}{|{\cal \hat{G}}|}  \sum_{x_k\in  {\cal \hat{G}}} \min_{y\in \mathcal M}\|x_k-y\|_2.
\]
\begin{figure}[t] %  figure placement: here, top, bottom, or page
   \includegraphics[width=\linewidth]{circle8n.eps} 
   \includegraphics[width=\linewidth]{sphere8.eps} 
   %\setcaptionwidth{6in}
   \vspace{-1cm}
   \caption{Margin and Hausdorff illustration for ridges obtained from {\it l}-SCRE and SCRE}
   \label{fig:circle6}
\end{figure}

@@

\section{Uniform Framework for Manifold Fitting}\label{Uniform_Framework}

>>>>

Today, the process of dealing with high-dimensional data requires the exponential consumption of computational resources, and is therefore costly but effective. Yet, even so, challenges remain, as the high-dimensional data tend to lie near a low-dimensional sub-manifold of the ambient space, which is a phenomenon termed ``the curse of dimensionality'' by \cite{fefferman2018fitting}.  Although, the geometric foundation of statistical inference in nonlinear regression was brought to attention of the statistical community as early as the 1980s by \cite{Kass1989}, the role of geometry was not primarily explored back then. At present, there are numerous works concerning manifold learning, with different areas of focus, such as dimension reduction focusing on approximating the embedding \citep{roweis2000nonlinear,zhang2004principal,donoho2003hessian} and manifold approximation via fitting the unknown manifold \citep{genovese2012manifold,chen2015asymptotic,genovese2014nonparametric}. In another seemingly-related direction, there have been works on statistical methods on manifolds over the past decades, centering on finding the mode of distribution \citep{huckemann2010intrinsic,chen2016nonparametric,chen2016comprehensive} and searching for the principal components of manifold \citep{hauberg2015principal,huckemann2006principal,hastie1989principal}.

There has also been some theoretical research into manifold learning using hypothesis testing. For instance, \cite{fefferman2018fitting} provided a condition and proved that the produced putative manifold ${\mathcal M}_o$'s Hausdorff distance to $\mathcal M$ is small, and its reach is not much smaller than that of $\mathcal M$. Mohammed and Narayanan proposed that the output manifold $\mathcal M$ can be defined by the set of points where the gradient of the approximate squared distance function (asdf) is orthogonal to the subspace spanned by the largest $D-d$ eigenvectors of the Hessian of the asdf \cite{mohammed2017manifold}.

@@

Manifold-fitting algorithms intend to recover the low-dimensional manifold in the ambient space by supposing that the observation data is generated from the manifold with some noise added. There are many classical algorithms to find the lower-dimensional manifold, such as the principal flow \cite{panaretos2014principal} and the local principal curves and surfaces based on kernel-density estimation \cite{ozertem2011locally}. The principal flow attempts to fit a curve moving along the maximal variation of the data, subject to a smoothness constraint. The solution to the principal flow is reduced to an ODE problem via the Euler-Lagrange method.  \cite{ozertem2011locally} redefined the principal curves and surfaces in terms of the gradient and Hessian of the probability density estimate. They also provided an algorithm based on the classical KDE. \cite{jung2012analysis} propose a general framework to fit the high-dimensional data with principal nested spheres. In addition, there is a large amount of literature about searching for hidden structure with an application in the point clouds, nebular hypothesis, earthquake locations, and object-oriented objects \cite{adams2011morse,davenport2010joint,klemela2009smoothing,book}.

@@

\begin{figure}[h] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=\linewidth]{field_demo.eps} 
   \caption{Illustration of the vector field. Left: using partial nearest samples Right: using all samples}
   \label{fig:example}
\end{figure}

The effect of improving the quality of $C(x)$ by neglecting samples with a large scale can be seen in the vector field in the above figure. The vector field is constructed from the dominant eigenvector of the weighted covariance matrix $C(x)$. In the left diagram, the set $S$ is the $k$-nearest neighborhood of $x$, and in the right diagram the set $S$ represents all the samples. From the figure, we know that by selecting the nearest samples we can get a weighted covariance matrix to approximate the vector field better.

Based on the above intuition, we show that constructing the {\it l}-SCRE by neglecting the far-away samples will help to reduce the point-wise bias of the derivatives. The reduced-bias phenomenon will further lead us to a ridge that is `closer' than that obtained by using all the samples.
 \subsection{Main Contributions}
Our contributions consist mainly of the following:
\begin{itemize}
\item[1.] We propose to use the nearest samples to construct a local version of the SCRE method, and in turn use that to construct the ridge-estimator function. The proposed {\it l}-SCRE  can obtain a ridge that is closer to the original manifold.
\item[2.] We provide the theoretical result of the component-wise bias and variance corresponding to the {\it l}-SCRE and prove that, under mild conditions, the local kernel function can indeed reduce the bias of derivatives more effectively than the KDE-based approaches.
\item[3.] By building the connection Hessian of the estimated density function with the empirical covariance matrix, we show the necessity of rank-one modification to obtain a stable tangent space estimator, which further improves the performance of the {\it l}-SCRE.
%\item[1.] We develop the connection between the two common methods (kernel-density estimation and tangent-space estimation) to define the fitted manifold.
%\item[2.] By analyzing the eigenspace of the Hessian matrix $H(x)$ of the kernel-density estimation function, we show that the normal space is not always generated by the principal eigenvectors of $H(x)$. To solve this problem, we split the $H(x)$ into two parts, $A(x)$ and $B(x)$. By estimating the normal space from $B(x)$, we can get a stable estimate of the normal space. The numerical experiment also demonstrates the effectiveness of our method.
%\item[3.] Our method can not only move an outlier $x_i$ onto the manifold, but also restrict the path of the convergence to be in the normal space of the intended targeting point $c(x)$.
%\item[4.] We also provide a faster algorithm to cope with the vanishing-gradient phenomenon for the kernel-density function.
\end{itemize}
Finally, we provide numerical examples to show the effectiveness of our strategy, and also show that the vector field corresponding to our method is smoother and more appropriate for the manifold-estimation problem.
\subsection{Notations}
The most frequently used symbols and notations are listed in Table \ref{symbols}.

\begin{table*}[h]\label{symbols}
\caption{Symbols and Notations}
\resizebox{14cm}{!}{
\begin{tabular}{c|c}\hline
Symbols & Meaning and Explanation\\\hline
$d$,$D$ & The dimension of the manifold and the dimension of the ambient space\\
${\rm E},{\rm Var}$ & The expectation and the variance operator\\
$\mathcal M,\hat{\mathcal G}$ & The true unknown manifold and the ridge estimator (discrete set)\\ 
$\mathcal M_{\hat{G}}$ & The projection of the ridge estimator $\hat{\mathcal G}$ onto $\mathcal M$\\
$p_n(x),g_n(x),H(x)$& The smooth density function, its gradient, and the Hessian Matrix\\
$\hat{p}_{r,h}(x), g_{\hat{p}_{r,h}}(x),H_{\hat{p}_{r,h}}(x)$& The local KDE estimator, its gradient, and the Hessian Matrix\\
 $\Pi,\Pi^c$ & The projection onto the top $d$ eigenvectors and its orthogonal complement\\
 $f$ & The nonlinear increasing, nonnegative, and concave function\\ 
 $J(x)$ & The semi-positive definite matrix, which is a key component of the Hessian\\  
 {$C(x)$, $H(x)$} &  {The weighted sample covariance matrix and the Hessian matrix at $x$}\\
 $\Pi_H$ & The projection matrix constructed by the top $d$ eigenvectors of $H$ \\ 
 ${\mathcal D}_r(x_0)$ & The $D$-dimensional ball centered at $x_0$ with a radius of $r$\\
 ${\cal N}(x, r)$ & The samples in the neighborhood of $x$ with a radius of $r$ \\
 \hline 
\end{tabular}}
\end{table*}

@@

We adopt the classical kernel-density estimation approach to handle our manifold-fitting problem, as the majority of previous works do. From the kernel-density perspective, the manifold can be viewed as points that satisfy some specific ridge conditions of the estimated density function.
\subsection{Ridge Definition}
There are two main definitions of a ridge, both of which involve describing the eigenspace of the Hessian and the gradient.
\begin{definition}\label{RR}
A point $x$ is on the $d$-dimensional ridge $R$ of its probability density function when the gradient $g(x)$ is orthogonal to at least $D-d$ eigenvectors of $H(x)$ and the corresponding $D-d$ eigenvalues are negative. \cite{ozertem2011locally}
\end{definition}
This definition gives us only a condition to test whether a point $x$ lies on the $d$-dimensional ridge. Given any $x$, it does not necessarily tell us how to pull it onto the $d$-dimensional ridge. There are $D-d$ eigenvectors of $H(x)$ that are orthogonal to the gradient $g(x)$, but it does not tell us the eigenvalues of $H(x)$ to which the eigenvectors correspond.

@@

\subsection{Ridge Estimation and Error Measurement}
In practice, we use the ridge $R$ as an approximation of the unknown manifold $\mathcal M$ numerically. However, the definition of a ridge in \eqref{ridgeDF} is arguably too descriptive and unrealistic. We can only judge whether a point resides on the ridge, but the definition provides an explicit continuous function to describe the ridge. In reality, starting from a random sample given in advance, we will need an algorithm to push the data points onto the ridge so that the resultant points satisfy the ridge definition. To be specific, what we obtain is a set of modified discrete samples, $\hat{\mathcal G}$, satisfying the ridge definition. To this end, we may further specify the ridge estimator $\hat{\cal G}$ in \eqref{Estimator} by
\[
\hat{\mathcal G} = {\rm Algorithm}(\{x_i, i=1:n\},  \hat{R})
\]
where the estimated $\hat{R}$ can be $\hat{R}_h$ or $\hat{R}_{r,h}$. Note that the resultant $\hat{\mathcal G}$ relies on the initial data points $\{x_i\}$, the estimated ridge $\hat{R}$, and the actual implementation (algorithm) involved in determining how to process the data $\{x_i\}$ by moving them onto the defined ridge. Obviously, we know $\hat{\mathcal G} \subset \hat{R}$.
\begin{figure}[htbp] %  figure placement: here, top, bottom, or page
   \centering
   \includegraphics[width=3.5in]{introduction_demo.eps} 
      \caption{Illustration of the $\hat{\mathcal G},\mathcal M, \mathcal M_{\hat{\mathcal G}}$ and the Hausdorff and average margin using the example of a toy circle. The blue diamond markers represent the observation points. The blue dots represent the computed points that satisfy the ridge condition, i.e. $\hat{\mathcal G}$. The red dots represent the projection of $\hat{\mathcal G}$ onto $\mathcal M$, denoted by $\cal M_{\hat{G}}$. }
   \label{fig:example}
\end{figure}

@@

To diminish the random factor in the distance measurement between $\hat{\mathcal G}$ and $\mathcal M$, we extract a subset from $\mathcal M$, ${\mathcal M}_{\hat{\mathcal G}}=\pi_{\mathcal M}(\hat{\mathcal G})$, and consider the Hausdorff distance between ${\hat{\mathcal G}}$ and ${\mathcal M}_{\hat{\mathcal G}}$. Since the projection $\pi_{\mathcal M}(\hat{\mathcal G})$ from $\hat{\mathcal G}$ to $\mathcal M_{\hat{G}}$ is onto, the Hausdorff distance equals the quasi-Hausdorff distance

@@

Lemma \ref{margin} builds the bridge between the geometric quantities and the analytic differential quantities. Through this lemma, we can easily convert the problem of searching for a `closer' ridge to the problem of searching for a `better' approximation of the Hessian and gradient. This conversion arguably give our work a stronger foundation and greater versatility.

@@

   \label{fig:example}

@@

\begin{example}
When we approximate $\hat{p}(x), g_{\hat{p}}(x), H_{\hat{p}}(x)$ using the kernel-density function as $\hat{p}_h(x), g_{\hat{p}_h}(x), H_{\hat{p}_h}(x)$, %where the subscript $n$ means the functions are related with the samples and the sample size $n$.
the pairwise error for the derivatives yields the following result:
\[
\begin{aligned}
& ( |g_{\hat{p}_h}(x)-g_p(x)|)_i = O(h^2) + O_p(\sqrt{\frac{1}{{{n h^{D+2}}}}}),\\
& ( |H_{\hat{p}_h}(x)-H_p(x)| )_{ij} = O(h^2) + O_p(\sqrt{\frac{1}{{{n h^{D+4}}}}}).
\end{aligned}
\]
The empirical-process theory gives the uniform error as
\[
\begin{aligned}
& \sup_x( |g_{\hat{p}_h}(x)-g_p(x)|)_i = O(h^2) + O_p(\sqrt{\frac{\log n}{{{n h^{D+2}}}}}),\\
& \sup_x( |H_{\hat{p}_h}(x)-H_p(x)| )_{ij} = O(h^2) + O_p(\sqrt{\frac{\log n}{{{n h^{D+4}}}}}).
\end{aligned}
\]
The error rates for the pairwise and uniform differ only in the stochastic term, with a scale of $\sqrt{\log n}$. \cite{chen2017tutorial,JMLR:v17:ariascastro16a, genovese2014nonparametric}.
\end{example}
%From \eqref{Hausmargin}, we can conclude that the Hausdorff distance of ${\mathcal G}$ and ${\mathcal M}_{\mathcal G}$ is controlled by the uniform approximation error of the Hessian and gradient, as $\sup_x( \|H_n(x)-H(x)\|_2+\|g_n(x)-g(x)\|_2$. The uniform derivatives rate is proved to be $L_{\infty}$, and the norm as , where the difference is the result of introducing $\log n$. This result is shown in \cite{JMLR:v17:ariascastro16a, genovese2014nonparametric}.
%Clearly, the Hausdorff distance between is the largest gap. As shown in \cite{genovese2014nonparametric}, the Hausdorff distance between the manifold and the ridge yields a uniform bound 
%${\rm Haus}({\mathcal G}, {\mathcal M}_{\mathcal G}) = O(h^2) + O_p((\frac{\log n}{{{n h^{d+4}}}})^{1/2})$. When $h={(\frac{\log n}{n}})^{1/(d+8)}$, the optimal minimum achieved is $O_p((\frac{\log n}{n})^{2/(d+8)})$. This optimum is the estimated convergent speed for the Hausdorff distance with the process of increasing the sample size $n$ and simultaneously decreasing $h$. However, in this paper, we mainly focus on how to get a closer ridge under the current observations, i.e. the number of samples is fixed.

@@

From Lemma \ref{improve_lemma}, we know that by reducing the bias with a scalar of $\ell$ we will have the optimal minimum solution reduced with a scalar of $\ell^{\frac{D+m}{D+m+4}}$. It should be noted that, with the increase in the dimension $D$, we will obtain a smaller $\ell^{\frac{D+m}{D+m+4}}$. Lemma \ref{improve_lemma} proves that we can improve the distance between the ridges by reducing the derivatives' bias corresponding to the density function. As before, the details can be found in the supplementary material.%appendix \ref{mimimum_proof}.

@@

\section{Numerical Experiments}\label{experiments}
In this section, we will show the effectiveness of our method compared with the classical subspace-constrained methods.
\subsection{Ridge Criteria}
The criterion for a point on a ridge is that it must satisfy the ridge condition. The ridge condition indicates that, when $x$ is on the ridge, the attraction vector $F(x)$ resides in the tangent space. The tangent space is usually obtained from the top d eigenvectors of the covariance matrix. Conversely, if the ridge condition is not met at $x$, we can also expect the angle between the attraction vector $F(x)$ and the estimated eigenspace to be large, i.e. a small cos value.

@@

   \begin{figure}[h] %  figure placement: here, top, bottom, or page
    \includegraphics[width=\linewidth]{vectorfield2.eps} 
    % \setcaptionwidth{6in}
    \caption{Illustration of the vector field corresponding to the eigenvector of the second eigenvalue with respect to three difference covariance matrices }
    \label{fig:vectorfield}
 \end{figure}
In Figure \eqref{fig:vectorfield}, we plot the vector field generated by the eigenvector corresponding to the smallest eigenvalue. This vector field is the geometric interpretation of the projection $\Pi^{\perp}$ in the two-dimensional space. Because the constrained subspace in the two-dimensional space is one-dimensional, the vector field coincides with the trajectory of the outliers moving onto the ridge. We can conclude from the above figures that the vector field corresponding to {\it l}-SCRE is more smooth and points approximately to the projection onto the manifold, while the vector field of SCRE is not well behaved for samples not close to the manifold.
\subsection{Synthetic Data Set}
First, we provide a one-dimensional ring example to show the effectiveness of our method compared with SCRE. Because of the simplicity of the circle structure, it is easy to compute the real projection onto it. We assume that our hidden manifold ${\mathcal M}$ is the one-dimensional circle embedded in a two-dimensional space with a radius of $1$ and center of $(0,0)$.

@@

For each point $\hat{x}$ belonging to ${\cal \hat{G}}$ or ${\cal \hat{G}}_\ell$, we can define the projection of $\hat{x}$ onto the real manifold $\mathcal M$ as $\pi_{\mathcal M}(\hat{x}) = \arg\min_{y\in \mathcal M} \|y-\hat{x}\|_2$. In the case of the special two-dimensional circle, the projection has an explicit form, $\pi_{\mathcal M}(\hat{x}) = \hat{x}/\|\hat{x}\|$. The average margin from $\cal \hat{G}$ or ${ \cal \hat{G}}_\ell$ to $\mathcal M$ is defined as
\[
{\rm Margin}({ \cal \hat{G}, \mathcal M}) =\frac{1}{|{\cal \hat{G}}|}  \sum_{x_k\in  {\cal \hat{G}}} \min_{y\in \mathcal M}\|x_k-y\|_2.
\]
\begin{figure}[t] %  figure placement: here, top, bottom, or page
   \includegraphics[width=\linewidth]{circle8.eps} 
   %\setcaptionwidth{6in}
   \vspace{-1cm}
   \caption{Margin and Hausdorff illustration for ridges obtained from {\it l}-SCRE and SCRE}
   \label{fig:circle6}
\end{figure}

@@

\section{Uniform Framework for Manifold Fitting}
